sdmTMB: an R package for fast, flexible, and user-friendly generalized linear mixed effects models with spatial and spatiotemporal random fields

### 2022-11-03 doc/paper-jss/paper-jss.Rmd
Fix Table label

\caption{Comparison of functionality between several \proglang{R} packages that can fit geostatistical GLMMs.}
\centering
\fontsize{8}{10}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{lllllll}
\label{tab:functionality}
\toprule
 & \pkg{sdmTMB} & \pkg{VAST} & \pkg{INLA}/\pkg{inlabru} & \pkg{mgcv} & \pkg{spBayes} & \pkg{spaMM}\\
\midrule
Time-varying coefficients & \checkmark & --$^1$ & \checkmark & \checkmark & \checkmark & --\\
Spatially varying coefficients (SVC) & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & --\\
GAMs$^2$ & \checkmark & -- & \checkmark & \checkmark & -- & --\\
Threshold covariates & \checkmark & -- & \checkmark$^3$ & -- & -- & --\\
Offsets & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
\addlinespace
Spatiotemporal fields & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark$^4$ & --\\
Spatial + spatiotemporal fields & \checkmark & \checkmark & \checkmark & \checkmark & -- & --\\
Anisotropy & \checkmark & \checkmark & -- & \checkmark & -- & --\\
Correlation barriers & \checkmark & \checkmark & \checkmark & \checkmark & -- & --\\
Separate range parameters for fields & \checkmark & -- & \checkmark & \checkmark & -- & --\\
Share range parameters across fields & \checkmark & \checkmark & \checkmark & -- & -- & --\\
SPDE-based & \checkmark & \checkmark & \checkmark & \checkmark$^5$ & --$^6$ & \checkmark\\
\addlinespace
NB1 distribution & \checkmark & -- & \checkmark & \checkmark & -- & \checkmark\\
NB2 distribution & \checkmark & \checkmark$^7$ & \checkmark & \checkmark & -- & \checkmark\\
Zero-truncated distributions & \checkmark & -- & \checkmark & -- & -- & \checkmark\\
Zero-inflated distributions & \checkmark & \checkmark & \checkmark & -- & -- & \checkmark\\
Tweedie distribution & \checkmark & \checkmark & \checkmark & \checkmark$^8$ & -- & --\\
Student-t distribution & \checkmark & -- & \checkmark & \checkmark & -- & --\\
Censored Poisson distribution & \checkmark & -- & \checkmark & -- & -- & --\\
log Gaussian Cox processes & --$^9$ & --$^9$ & \checkmark & --$^9$ & --$^9$ & --$^9$\\
Multivariate responses & -- & \checkmark & \checkmark & -- & \checkmark & \checkmark\\
Built-in delta/hurdle models & \checkmark & \checkmark & \checkmark & --$^{10}$ & -- & \checkmark\\
Poisson-link delta model & \checkmark & \checkmark & \checkmark & -- & -- & --\\
Likelihood weights & \checkmark & -- & \checkmark & \checkmark & \checkmark & \checkmark\\
Maximum/marginal likelihood & \checkmark & \checkmark & -- & \checkmark & -- & --\\
\addlinespace
Bayesian/optionally Bayesian & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & --\\
Priors/penalties & \checkmark & -- & \checkmark & -- & \checkmark & --\\
Matern PC priors & \checkmark & -- & \checkmark & -- & -- & --\\
\addlinespace
Spatial (or spatial dynamic) factor analysis & -- & \checkmark & -- & -- & -- & --\\
Empirical Orthogonal Function (EOF) analysis & -- & \checkmark & -- & -- & -- & --\\
Built-in area-weighted index standardization & \checkmark & \checkmark & -- & -- & -- & --\\
Built-in cross-validation & \checkmark & -- & -- & -- & -- & --\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item $^1$Technically possible but non-trivial. $^2$Penalized smoother GAMs that determine `wiggliness'. $^3$\pkg{inlabru} but not \pkg{INLA}. $^4$Spatiotemporal fields as random walk only. $^5$SPDE approach as in Miller et al. (2019). $^6$Does have predictive process knots. $^7$Zero-inflated NB2 only. $^8$Tweedie power parameter fixed for \texttt{mgcv::gamm()}. $^9$Possible as log-linked Poisson GLMM with aggregated data. $^{10}$Hurdle models possible by fitting components separately.
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

```{r child="model-description.Rmd", echo=FALSE, cache=FALSE, eval=TRUE}
```

```{r child="pcod.Rmd", echo=FALSE, cache=FALSE, eval=TRUE}
```

```{r child="owls.Rmd", echo=FALSE, cache=FALSE, eval=TRUE}
```

```{r child="inla-comparison.Rmd", echo=FALSE, cache=FALSE, eval=TRUE}
```

```{r child="timing.Rmd", echo=FALSE, cache=FALSE, eval=TRUE}
```

>>>>

\caption{Comparison of functionality between several \proglang{R} packages that can fit geostatistical GLMMs.}
\centering
\fontsize{8}{10}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{lllllll}
\toprule
 & \pkg{sdmTMB} & \pkg{VAST} & \pkg{INLA}/\pkg{inlabru} & \pkg{mgcv} & \pkg{spBayes} & \pkg{spaMM}\\
\midrule
Time-varying coefficients & \checkmark & --$^1$ & \checkmark & \checkmark & \checkmark & --\\
Spatially varying coefficients (SVC) & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & --\\
GAMs$^2$ & \checkmark & -- & \checkmark & \checkmark & -- & --\\
Threshold covariates & \checkmark & -- & \checkmark$^3$ & -- & -- & --\\
Offsets & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
\addlinespace
Spatiotemporal fields & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark$^4$ & --\\
Spatial + spatiotemporal fields & \checkmark & \checkmark & \checkmark & \checkmark & -- & --\\
Anisotropy & \checkmark & \checkmark & -- & \checkmark & -- & --\\
Correlation barriers & \checkmark & \checkmark & \checkmark & \checkmark & -- & --\\
Separate range parameters for fields & \checkmark & -- & \checkmark & \checkmark & -- & --\\
Share range parameters across fields & \checkmark & \checkmark & \checkmark & -- & -- & --\\
SPDE-based & \checkmark & \checkmark & \checkmark & \checkmark$^5$ & --$^6$ & \checkmark\\
\addlinespace
NB1 distribution & \checkmark & -- & \checkmark & \checkmark & -- & \checkmark\\
NB2 distribution & \checkmark & \checkmark$^7$ & \checkmark & \checkmark & -- & \checkmark\\
Zero-truncated distributions & \checkmark & -- & \checkmark & -- & -- & \checkmark\\
Zero-inflated distributions & \checkmark & \checkmark & \checkmark & -- & -- & \checkmark\\
Tweedie distribution & \checkmark & \checkmark & \checkmark & \checkmark$^8$ & -- & --\\
Student-t distribution & \checkmark & -- & \checkmark & \checkmark & -- & --\\
Censored Poisson distribution & \checkmark & -- & \checkmark & -- & -- & --\\
log Gaussian Cox processes & --$^9$ & --$^9$ & \checkmark & --$^9$ & --$^9$ & --$^9$\\
Multivariate responses & -- & \checkmark & \checkmark & -- & \checkmark & \checkmark\\
Built-in delta/hurdle models & \checkmark & \checkmark & \checkmark & --$^{10}$ & -- & \checkmark\\
Poisson-link delta model & \checkmark & \checkmark & \checkmark & -- & -- & --\\
Likelihood weights & \checkmark & -- & \checkmark & \checkmark & \checkmark & \checkmark\\
Maximum/marginal likelihood & \checkmark & \checkmark & -- & \checkmark & -- & --\\
\addlinespace
Bayesian/optionally Bayesian & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & --\\
Priors/penalties & \checkmark & -- & \checkmark & -- & \checkmark & --\\
Matern PC priors & \checkmark & -- & \checkmark & -- & -- & --\\
\addlinespace
Spatial (or spatial dynamic) factor analysis & -- & \checkmark & -- & -- & -- & --\\
Empirical Orthogonal Function (EOF) analysis & -- & \checkmark & -- & -- & -- & --\\
Built-in area-weighted index standardization & \checkmark & \checkmark & -- & -- & -- & --\\
Built-in cross-validation & \checkmark & -- & -- & -- & -- & --\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item $^1$Technically possible but non-trivial. $^2$Penalized smoother GAMs that determine `wiggliness'. $^3$\pkg{inlabru} but not \pkg{INLA}. $^4$Spatiotemporal fields as random walk only. $^5$SPDE approach as in Miller et al. (2019). $^6$Does have predictive process knots. $^7$Zero-inflated NB2 only. $^8$Tweedie power parameter fixed for \texttt{mgcv::gamm()}. $^9$Possible as log-linked Poisson GLMM with aggregated data. $^{10}$Hurdle models possible by fitting components separately.
\end{tablenotes}
\end{threeparttable}
\label{tab:functionality}
\end{table}

@@

```{r child="model-description.Rmd", eval=TRUE}
```

```{r child="pcod.Rmd", eval=TRUE}
```

```{r child="owls.Rmd", eval=TRUE}
```

```{r child="inla-comparison.Rmd", eval=TRUE}
```

```{r child="timing.Rmd", eval=TRUE}
```
